---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rTriathlon

<!-- badges: start -->

<!-- badges: end -->

rTriathlon is a client for <https://www.triathlon.org/>.

Resources regarding Triathlon qualifications:

-   <https://www.teamusa.org/-/media/USA_Triathlon/PDF/Elite-International/2023/TRI_2024_OLY_ATH_FINAL-With-Signature-NEW.pdf?la=en&hash=327A5F72C8CA11AB41CFEEC84B8D6E1A5CE544C3>

-   <https://en.wikipedia.org/wiki/Triathlon_at_the_2024_Summer_Olympics_%E2%80%93_Qualification>

-   <https://olympics.com/en/news/pathway-to-paris-triathlon-qualification-system-explained>

-   <https://www.triathlon.org/news/article/ioc_approves_the_olympic_qualification_criteria_for_paris_2024_olympics>

-   <https://www.triathlon.org/uploads/docs/Paris2024_Olympic_Qualification_Principles.pdf?fbclid=IwAR2daD8STkBG8uBQfAYEXAGFArwEGFUT5xABpDaCYxvp_EPpRpVZ2P-FFV0>

## Installation

You can install the development version of rTriathlon from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tyler-widdison/rTriathlon")
```

## Example

Get a list of triathlon events (this date range happens to be the Olympic qualification range):

```{r}
library(rTriathlon)
event_list <- tri_get_event_list('2022-05-27', '2024-05-27')
nrow(event_list)
```

Get details for a specific event:

```{r}
event_id <- event_list$event_id[1:5]
prog_list <- tri_program_list_get(event_id)
prog_list %>% dplyr::select(prog_id, event_id, prog_name) %>% head
```

We can then get results from events:

```{r}
tri_results_list_get(174221, 583767)
```

If we only want to consider USA athletes Olympic qualifying events per <https://triathlon.org/rankings>

```{r}
# Get events in qualifying current date range
oly_events <- event_list %>%
    dplyr::mutate(event_date = as.Date(event_date)) %>% 
    tidyr::unnest(event_categories, keep_empty = T) %>% 
    dplyr::filter(cat_id %in% c(351, 348, 345, 349, 341, 340) & event_id != '159913') %>% # filtering out 2022 GBR has used this to qualify already
    dplyr::mutate(keep = ifelse(grepl('Continental', cat_name) & !grepl('America', event_title), 'toss', 'keep')) %>% 
    dplyr::filter(keep == 'keep') %>%
    dplyr::mutate(Year = lubridate::year(event_date),
           Month = lubridate::month(event_date, label = F, abbr = F),
           Day = lubridate::wday(event_date, label = T),
           mday = lubridate::mday(event_date),
           Month_week = (5 + lubridate::day(event_date) + lubridate::wday(lubridate::floor_date(event_date, 'month'))) %/% 7)

oly_events %>% head
```

We can then look at the results for the qualifying `oly_events` events.

```{r}
prog_list <- tri_program_list_get(oly_events$event_id)
prog_list <- prog_list %>% dplyr::filter(is_race == T & results == T)

# Function to apply tri_results to multiple events/catagories
get_results <- function(event_nos, prog_nos) {
  results_list <- lapply(seq_along(event_nos), function(i) {
    tryCatch(
      tri_results_list_get(event_no = event_nos[i], prog_no = prog_nos[i]),
      error = function(e) {
        message(paste0("Error retrieving results for event ", event_nos[i], " and program ", prog_nos[i], ": ", e$message))
        return(NULL)
      }
    )
  })
  results_df <- do.call(plyr::rbind.fill, results_list)
  return(results_df)
}

data <- get_results(prog_list$event_id, prog_list$prog_id)

# Join the results back to the oly_events data for event data
df <- data %>% 
  dplyr::left_join(oly_events %>% 
              dplyr::select(event_id, cat_id, event_title, event_date, cat_name), 
            by = c('event_id'))

df %>% head
```

We can check out how each NOC is fairing thus far:

```{r, echo=FALSE}
df %>% 
    dplyr::filter(!position %in% c('DNF', 'DNS', 'DSQ', 'LAP', 'NC') & 
           !is.na(position) & athlete_gender != '') %>% 
    dplyr::mutate(position = as.numeric(position)) %>% 
    dplyr::group_by(event_id, event_title, athlete_gender, cat_name, athlete_noc, position) %>% 
    dplyr::count() %>% 
    dplyr::ungroup() %>% 
    dplyr::filter(n > 2) %>% 
    ggplot2::ggplot(ggplot2::aes(stats::reorder(athlete_noc, position), position, fill = cat_name)) + 
    ggplot2::geom_boxplot() + 
    ggplot2::theme_bw() + 
    ggplot2::labs(y = 'Average Athlete NOC Finish',
         title = 'Combined M/W average finishes during Olympic \nqualification period with more than 4 total competitions',
         x = '',
         subtitle = 'Data filters out events less then 2') + 
  ggplot2::coord_flip()
```

### Version control

Currently in a developmental state
